---
title: "Covid91 vaccine study Final2022"
author: "Precious Calderon"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="final_Data.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
options(digits = 1)
```

```{r}
Males <- filter(FinalData, sex=="M")
Females <- filter(FinalData, sex == "F")
LGBTQ <- filter(FinalData, LGBTQ == "gay")
DrugUser <- filter(FinalData, DrugUser == "yes")
```



# Introduction

A new  preventative vaccine (covid91) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

# Method

The original data set, FinalDat, contains all character variables that are needed to be converted into factor variables in order to analyze the results of the study. This new data set is named "FinalData" which will then be divided into the four subgroups mentioned in the introduction. These subgroups are males, females, people who identify as LGBTQ, and drug users. Part of each sub group has taken the vaccine while the remaining part has taken a placebo, so the effects of these two types of treatments would be analyzed. The vaccine would be considered efficacious if it has a statistically significant effect on people who have taken it. In other words, a significant number of people in a sub group did not get covid-19 as a result of getting vaccinates. Lastly, certain recommendations will be made regarding whether each groups should consider taking or not taking the covid-19 vaccine based on the analyzed results. 


## Males

### Descriptive Results

#### Graphical Descriptive Results

```{r}
dd2 <- Males %>% group_by(infected, treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

The bar plots above are graphical descriptive summaries of the effect of the two treatments (covid 19 vaccine and placebo) on males. At first glance, the vaccine appears to be working among males as the two plots show that more males who took the placebo contracted the disease. 



#### Numerical Descriptive Results

```{r}
table1 <- xtabs(~infected + treatment, data=Males)
rowPerc(table1)
colPerc(table1)
```
The most important info: among males who got covid19, 67.78% of them took the placebo and only 32.22% took the vaccine. Majority of the males who got covid took placebo, therfore the vaccine seems to be working among males. 


### Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
malefisher <- fisher.test(table1)
malefisher
```


Observed counts table shows the actual number of males who got covid and those who didn't. Also, p-value is small meaning there is a statistical difference in the effect of the placebo and the vaccine among males in preventing covid infection. Males who took the placebo are about 2 times more likely to contract covid than those who took the vaccine based on the odds ratio value. 

### Efficacy of Vaccine for Males

```{r}

EfficacyMales <- ((malefisher$estimate/(1+malefisher$estimate)*100) )
malefisherlow <- malefisher$conf.int[1]
malefisherhigh <- malefisher$conf.int[2]
EfficacyMalesLow <- (malefisherlow/(1+malefisherlow)*100)
EfficacyMalesHigh <- (malefisherhigh/(1+malefisherhigh)*100)
```
The estimated efficacy of the vaccine among males is `r EfficacyMales`%. The 95% confidence interval of the vaccine efficacy among males is from `r EfficacyMalesLow` to `r EfficacyMalesHigh`. 

## Females

### Descriptive Results

#### Graphical Descriptive Results

#### Numerical Descriptive Results

### Inferential Results

### Efficacy of Vaccine for Males



## LGBTQ

### Descriptive Results

#### Graphical Descriptive Results

#### Numerical Descriptive Results

### Inferential Results

### Efficacy of Vaccine for Males



## Druggies

### Descriptive Results

#### Graphical Descriptive Results

#### Numerical Descriptive Results

### Inferential Results

### Efficacy of Vaccine for Males


## Discussion and Conclusion

Note: paste values of efficacy of vaccine for each group. 
