---
title: "Covid91 vaccine study Final2022"
author: "Precious Calderon"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="final_Data.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
options(digits = 1)
```

```{r}
Males <- filter(FinalData, sex=="M")
Females <- filter(FinalData, sex == "F")
LGBTQ <- filter(FinalData, LGBTQ == "gay")
DrugUser <- filter(FinalData, DrugUser == "yes")
```



# Introduction

A new  preventative vaccine (covid91) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of sub populations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

# Method

The original data set, FinalDat, contains all character variables that are needed to be converted into factor variables in order to analyze the results of the study. This new data set is named "FinalData" which will then be divided into the four subgroups mentioned in the introduction. These subgroups are males, females, people who identify as LGBTQ, and drug users. Part of each sub group has taken the vaccine while the remaining part has taken a placebo, so the effects of these two types of treatments would be analyzed. The vaccine would be considered efficacious if it has a statistically significant effect on people who have taken it. In other words, the calculated efficacy value for the vaccine must be greater than 50%. Lastly, certain recommendations will be made regarding whether each groups should consider taking or not taking the covid-19 vaccine based on the analyzed results. 


## Males

### Descriptive Results

#### Graphical Descriptive Results

```{r}
dd2 <- Males %>% group_by(infected, treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

The two plots above show the graphical summary of the relationship between the type of treatment (placebo and vaccine) and the contraction of covid-19 specifically among males.     

The first bar plot shows the estimated counts of males who did and did not get covid and whether they took the vaccine or the placebo. At first glance, it can be seen that the majority of the male population did not contract covid-19 regardless of the type of treatment they received. However, looking closely, the bar representing males who took the placebo and contracted covid-19 is larger compared to the bar representing males who took the vaccine and contracted covid-19. This may suggest that the vaccine is working among males given that there is less counts of infection among males who took the vaccine.   

Meanwhile, the second bar plot shows the estimated percentages of covid-19 infection among males who took the two types of treatment. Among males who contracted covid-19, about 65% of them took the placebo, and 35% got the vaccine. This means that infection among males is largely made up males who took the placebo. Furthermore, among males who did not contract the virus, there apears to be a 50:50 ratio of males who took the placebo and got the vaccine.   

Based on these two plots, the vaccine appears to be working among males as they show that more males who took the placebo contracted covid-19 than males who got the vaccine.   
  
  
  
#### Numerical Descriptive Results

```{r}
table1 <- xtabs(~infected + treatment, data=Males)
rowPerc(table1)
colPerc(table1)
```
The tables above show the actual values of the percentages of males who did and did not contract covid-19 based on the treatment they received (placebo and vaccine). 

The most important piece of information from these two tables is seen from on the first table. According to it, among males who contracted covid-19, 68% of them took the placebo and only 32% took the vaccine. This reinforces the previous observation that majority of males who got covid-19 took placebo, and therefore, the vaccine seems to be working among males in preventing the contraction of covid-19. 


### Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
malefisher <- fisher.test(table1)
malefisher
```

To further test whether the vaccine is efficacious among males, chi-squares test is conducted along with fisher's exact test.   
  
Before going into the results of this test, notice that the observed counts table shows the actual number of males who did and did not get covid. There are 244 males who took the placebo and got covid-19 while there are only 116 males who got the vaccine and got covid-19.   
  
The chi-squared test provided a small p-value of 0 which means that results expected by the null hypothesis are unlikely to happen by chance. This also implies there is a correlation between the type of treatment and contraction of covid-19 among males, or that there is a statistical difference in the effect of the placebo and the vaccine among males in preventing covid-19 infection.  Moreover, according to the odds ratio provided in the fisher's exact test, males who took the placebo are `r malefisher$estimate` times more likely to contract covid-19 than those who took the vaccine.  
  
  
### Efficacy of Vaccine for Males

```{r}

EfficacyMales <- ((malefisher$estimate/(1+malefisher$estimate)*100) )
malefisherlow <- malefisher$conf.int[1]
malefisherhigh <- malefisher$conf.int[2]
EfficacyMalesLow <- (malefisherlow/(1+malefisherlow)*100)
EfficacyMalesHigh <- (malefisherhigh/(1+malefisherhigh)*100)
```

To be able to conclude that the vaccine is indeed efficacious among males, the calculated value of efficacy must be greater than 50%. Otherwise, the vaccine will be deemed ineffective and will not be recommended for administration for this specific sub group.  
  
The estimated efficacy of the vaccine among males is `r EfficacyMales`%. The 95% confidence interval of the vaccine efficacy among males is from `r EfficacyMalesLow` to `r EfficacyMalesHigh`. Therefore, these results conclude that the vaccine is efficacious among males, so it's recommended that members of this population take the vaccine to prevent contracting covid-19.   

## Females

### Descriptive Results
  
#### Graphical Descriptive Results

```{r}
dd3 <- Females %>% group_by(infected, treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd3,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd3,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```
  
The two plots above show the graphical summary of the relationship between the type of treatment (placebo and vaccine) and the contraction of covid-19 specifically among females.  
  
The first bar plot shows the estimated counts of females who did and did not get covid-19 and whether they took the vaccine or the placebo. At first glance, it can be seen that the majority of the female population did not contract covid-19 regardless of the type of treatment they received. However, looking closely, the bar representing females who took the placebo and contracted covid-19 is larger compared to the bar representing females who took the vaccine and contracted covid-19. This may suggest that the vaccine is working among females given that there is less counts of infection among females who took the vaccine.  

Meanwhile, the second bar plot shows the estimated percentages of covid-19 infection among females who took the two types of treatment. Among females who contracted covid-19, close to 60% of them took the placebo, and about 40% got the vaccine. This means that infection among females is largely made up females who took the placebo. Furthermore, among females who did not contract the virus, there appears to be a 50:50 ratio of females who took the placebo and got the vaccine.  
  
Based on these two plots, the vaccine appears to be working among females as they show that more females who took the placebo contracted covid-19 than females who got the vaccine.  
  
  
  
#### Numerical Descriptive Results

```{r}
table2 <- xtabs(~infected + treatment, data=Females)
rowPerc(table2)
colPerc(table2)
```
  
The tables above show the actual values of the percentages of females who did and did not contract covid-19 based on the treatment they received (placebo and vaccine).

The most important piece of information from these two tables is seen from on the first table. According to it, among females who contracted covid-19, 59% of them took the placebo and only 41% took the vaccine. This reinforces the previous observation that majority of females who got covid-19 took placebo, and therefore, the vaccine seems to be working among females in preventing the contraction of covid-19.
  
  
### Inferential Results
  
```{r}
chisq.test(table2)
chisqtestGC(table2)
femalefisher <- fisher.test(table2)
femalefisher
```
  
To further test whether the vaccine is efficacious among females, chi-squares test is conducted along with fisher’s exact test.  
  
Before going into the results of this test, notice that the observed counts table shows the actual number of females who did and did not get covid. There are 212 females who took the placebo and got covid-19 while there are 148 females who got the vaccine and got covid-19.

The chi-squared test provided a small p-value of 0.001 which means that results expected by the null hypothesis are unlikely to happen by chance. This also implies there is a correlation between the type of treatment and contraction of covid-19 among females, or that there is a statistical difference in the effect of the placebo and the vaccine among females in preventing covid-19 infection. Moreover, according to the odds ratio provided in the fisher’s exact test, females who took the placebo are `r femalefisher$estimate` times more likely to contract covid-19 than those who took the vaccine.

### Efficacy of Vaccine for Females
  
```{r}

EfficacyFemales <- ((femalefisher$estimate/(1+femalefisher$estimate)*100) )
femalefisherlow <- femalefisher$conf.int[1]
femalefisherhigh <- femalefisher$conf.int[2]
EfficacyFemalesLow <- (femalefisherlow/(1+femalefisherlow)*100)
EfficacyFemalesHigh <- (femalefisherhigh/(1+femalefisherhigh)*100)
```

To be able to conclude that the vaccine is indeed efficacious among females, the calculated value of efficacy must be greater than 50%. Otherwise, the vaccine will be deemed ineffective and will not be recommended for administration for this specific sub group.  
  
The estimated efficacy of the vaccine among females is `r EfficacyFemales`%. The 95% confidence interval of the vaccine efficacy among females is from `r EfficacyFemalesLow` to `r EfficacyFemalesHigh`. Therefore, these results conclude that the vaccine is efficacious among females, so it's recommended that members of this population take the vaccine to prevent contracting covid-19.   

## LGBTQ

### Descriptive Results
  
#### Graphical Descriptive Results

```{r}
dd4 <- LGBTQ %>% group_by(infected, treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd4,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd4,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

The two plots above show the graphical summary of the relationship between the type of treatment (placebo and vaccine) and the contraction of covid-19 specifically among people who identify as LGBTQ.   
  
The first bar plot shows the estimated counts of people who identify as LGBTQ who did and did not get covid-19 and whether they took the vaccine or the placebo. At first glance, it can be seen that the majority of the LGBTQ population did not contract covid-19 regardless of the type of treatment they received. However, looking closely, the bar representing LGBTQ members who took the placebo and contracted covid-19 is smaller compared to the bar representing LGBTQ members who took the vaccine and contracted covid-19. This may suggest that the vaccine is not working among the LGBTQ community given that there is more counts of infection among members who took the vaccine.  
  
Meanwhile, the second bar plot shows the estimated percentages of covid-19 infection among LGBTQ members who took the two types of treatment. Among people who identify as LGBTQ who contracted covid-19, only about 15% of them took the placebo, and about 85% got the vaccine. This means that infection among members of LBGTQ community is largely made up people who took the vaccine. Furthermore, among LGBTQ members who did not contract the virus, there appears to be a 50:50 ratio of LGBTQ people who took the placebo and got the vaccine.

Based on these two plots, the vaccine appears to be ineffective among members of the LBGTQ community as they show that more LGBTQ members who took the vaccine contracted covid-19 than those who got the placebo.


#### Numerical Descriptive Results

```{r}
table3 <- xtabs(~infected + treatment, data=LGBTQ)
rowPerc(table3)
colPerc(table3)
```
  
The tables above show the actual values of the percentages of people who identify as LGBTQ who did and did not contract covid-19 based on the treatment they received (placebo and vaccine).  
  
The most important piece of information from these two tables is seen from on the first table. According to it, among LGBTQ members who contracted covid-19, a large 87% of them took the vaccine and only 13% took the placebo This reinforces the previous observation that majority of LGBTQ members who got covid-19 took the vaccine, and therefore, the it does not seem to be working among the LGBTQ community in preventing the contraction of covid-19.  
  
  
### Inferential Results
  
```{r}
chisq.test(table3)
chisqtestGC(table3)
lgbtqfisher <- fisher.test(table3)
lgbtqfisher
```




### Efficacy of Vaccine for LGBTQ
  
```{r}

EfficacyLGBTQ <- ((lgbtqfisher$estimate/(1+lgbtqfisher$estimate)*100) )
lgbtqfisherlow <- lgbtqfisher$conf.int[1]
lgbtqfisherhigh <- lgbtqfisher$conf.int[2]
EfficacyLGBTQLow <- (lgbtqfisherlow/(1+lgbtqfisherlow)*100)
EfficacyLGBTQHigh <- (lgbtqfisherhigh/(1+lgbtqfisherhigh)*100)
```
  
To be able to conclude that the vaccine is efficacious among LGBTQ members, the calculated value of efficacy must be greater than 50%. Otherwise, the vaccine will be deemed ineffective and will not be recommended for administration for this specific sub group.  
  
The estimated efficacy of the vaccine among LGBTQ members is `r EfficacyLGBTQ`%. The 95% confidence interval of the vaccine efficacy among LGBTQ members is from `r EfficacyLGBTQLow` to `r EfficacyLGBTQHigh`. Therefore, these results conclude that the vaccine is inefficacious among people who identify as LGBTQ, so it is not recommended that members of this population take the vaccine to prevent contracting covid-19.   


## Drug Users

### Descriptive Results

#### Graphical Descriptive Results
  
```{r}
dd5 <- DrugUser %>% group_by(infected, treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd5,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd5,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

#### Numerical Descriptive Results

```{r}
table4 <- xtabs(~infected + treatment, data=DrugUser)
rowPerc(table4)
colPerc(table4)
```

### Inferential Results
  
```{r}
chisq.test(table4)
chisqtestGC(table4)
drugfisher <- fisher.test(table4)
drugfisher
```

### Efficacy of Vaccine for Males


## Discussion and Conclusion

Note: paste values of efficacy of vaccine for each group. 
